### 2016-05-25 paper/icfp16-cameraready/evaluation.tex
clarify size of benchmarks

\paragraph{Benchmarks}
We answer these questions on two sets of ill-typed programs, \ie
programs that were rejected by the \ocaml\ compiler because of a
type error.
%
The first dataset comes from the Spring 2014 undergraduate Programming
Languages course at UC San Diego.
%
We recorded each interaction with the \ocaml\ top-level system over the
course of the first three assignments (IRB  
% \# hidden for blind review),
\#140608),
from which we extracted \ucsdsize\ distinct, ill-typed \ocaml programs.
%
The second dataset --- widely used in the literature --- comes from a
graduate-level course at the University of Washington~\cite{Lerner2006-pj},
from which we extracted 284 ill-typed programs.
%
Both datasets include programs that demonstrate a variety of functional
programming idioms, including (tail) recursive functions, higher-order
functions, polymorphic data types, and expression evaluators.

\subsection{Witness Coverage}
\label{sec:eval:witness-coverage}
%
We ran our search algorithm on each program for 1,000 iterations, with
the entry point set to the function that \ocaml\ had identified as
containing a type error.
%
Due to the possibility of non-termination we set a timeout of 10 minutes
per program.
%
% Due to the possibility of non-termination we set a limit on the number
% of reductions to perform, increasing in 500-step increments from 500
% steps to 3,000 steps total.
%
We also added a na\"ive check for infinite recursion; at each recursive
function call we check whether the new arguments are identical to the
current arguments.
%
If so, the function cannot possibly terminate and we report an error.
%
While not a \emph{type error}, infinite recursion is still a clear bug
in the program, and thus valuable feedback for the user.

@@

\paragraph{Random Generation}
Random test generation has difficulty generating highly constrained
values, \eg red-black trees or a pair of equal integers. If the type
error is hidden behind a complex branch condition \nanomaly may not be
able to trigger it. Exhaustive testing and dynamic-symbolic execution
can address this short-coming by performing an exhaustive search for
inputs (\emph{resp}. paths through the program). As our experiments
show, however, novice programs do not appear to require more advanced
search techniques, likely because the novice programs tend to be simple.

@@

\paragraph{Trace Explosion}
Though the average complexity of our generated traces is low in terms of
jumps, there are some extreme outliers. We cannot reasonably expect a
novice user to explore a trace containing 50+ terms and draw a
conclusion about which pieces contributed to the bug in their
program. Enhancing our visualization to slice out program paths relevant
to specific values~\cite{Perera2012-dy},
would likely help alleviate this issue, allowing users to
highlight a confusing value and ask: ``Where did this come from?''

>>>>

\paragraph{Benchmarks}
We answer these questions on two sets of ill-typed programs, \ie\
programs that were rejected by the \ocaml\ compiler because of a
type error.
%
The first dataset comes from the Spring 2014 undergraduate Programming
Languages course at UC San Diego.
%
We recorded each interaction with the \ocaml\ top-level system over the
course of the first three assignments (IRB  
% \# hidden for blind review),
\#140608),
from which we extracted \ucsdsize\ distinct, ill-typed \ocaml\ programs.
%
The second dataset --- widely used in the literature --- comes from a
graduate-level course at the University of Washington~\cite{Lerner2006-pj},
from which we extracted 284 ill-typed programs.
%
Both datasets contain relatively small programs, the largest being 348
SLoC; however, they demonstrate a variety of functional programming
idioms including (tail) recursive functions, higher-order functions,
polymorphic data types, and expression evaluators.

\subsection{Witness Coverage}
\label{sec:eval:witness-coverage}
%
We ran our search algorithm on each program for 1,000 iterations, with
the entry point set to the function that \ocaml\ had identified as
containing a type error.
%
Due to the possibility of non-termination we set a timeout of 10 minutes
per program.
%
% Due to the possibility of non-termination we set a limit on the number
% of reductions to perform, increasing in 500-step increments from 500
% steps to 3,000 steps total.
%
We also added a na{\"\i}ve check for infinite recursion; at each recursive
function call we check whether the new arguments are identical to the
current arguments.
%
If so, the function cannot possibly terminate and we report an error.
%
While not a \emph{type error}, infinite recursion is still a clear bug
in the program, and thus valuable feedback for the user.

@@

\paragraph{Random Generation}
Random test generation has difficulty generating highly constrained
values, \eg red-black trees or a pair of equal integers. If the type
error is hidden behind a complex branch condition \nanomaly may not be
able to trigger it. Exhaustive testing and dynamic-symbolic execution
can address this short-coming by performing an exhaustive search for
inputs (\emph{resp}.\ paths through the program). As our experiments
show, however, novice programs do not appear to require more advanced
search techniques, likely because the novice programs tend to be simple.

@@

\paragraph{Trace Explosion}
Though the average complexity of our generated traces is low in terms of
jumps, there are some extreme outliers.
%
We cannot reasonably expect a novice user to explore a trace containing
50+ terms and draw a conclusion about which pieces contributed to the
bug in their program.
%
Enhancing our visualization to slice out program paths relevant to
specific values~\cite{Perera2012-dy}, would likely help alleviate this
issue, allowing users to highlight a confusing value and ask: ``Where
did this come from?''
